---
title: "Rthritis"
author: "the best group EVER"
date: '2022-05-06'
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE, 
                      dev.args=list(bg="transparent"))
```

```{r libraries, echo = TRUE}
library("tidyverse")
```

## 1. Introduction

Why we wanted to look at arthritis, what we wanted to find out, how we were going to do it (in very broad terms), bla bla bla bla.

## 2. Materials and methods

Explain the dataset and articles
Explain the order we did things

## 3. Tidy datasets

```{r cars, echo = TRUE}
summary(cars)
```

## 4. Results 1: Exploration of Metadata (Klara)

```{r}
plot(pressure)
```

## 5. Results 2: PCA and clusters (Klara) 

```{r}
plot(pressure)
```
  
## 6. Results 3: Differential expression (Mads)

```{r}

```

## 7. Results 4: Heatmaps (Javi)

```{r}
plot(pressure)
```

## 8. Results 5: Box plots (Javi)

```{r}
plot(pressure)
```

## 9. Results 6: Model (Angeliki)

```{r}
plot(pressure)
```

## 10. Discussion + conclusion (Angeliki)

```{r}
plot(pressure)
```

## Goodbye- slide



