---
title: "Rthritis"
author: "the best group EVER"
date: '2022-05-06'
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE, 
                      dev.args=list(bg="transparent"))
```

## 1. Introduction

Why we wanted to look at arthritis, what we wanted to find out, how we were going to do it (in very broad terms), bla bla bla bla.

## 2. Materials and methods

Explain the dataset and articles
Explain the order we did things

## 3. Result 1: Tidy datasets: Before vs. after.

```{r cars, echo = TRUE}
summary(cars)
```

## 4. Result 2: First look at data: Patient distribution. (Daniel)

```{r}
plot(pressure)
```

## 5. Result 3: Exploration: PCA plots and clustering (Klara)

```{r}
plot(pressure)
```
  
## 6. Result 4: Differential expression: Log2fold changes (Mads)

```{r}
plot(pressure)
```

## 7. Result 5: Heatmaps (Javi)

```{r}
plot(pressure)
```

## 8. Result 6: Modelling: Logistic regression (Angiliki)

```{r}
plot(pressure)
```

## 9. Extra ??

```{r}
plot(pressure)
```

## 10. Discussion + conclusion

```{r}
plot(pressure)
```

## Goodbye- slide



